# Genus phyloseq

All taxa levels from phyla to species to ASVs can be used with MaAsLin3.
In fact, in your own research I would recommend carrying out MaAsLin3 analysis on multiple taxa levels.

For speed and to produce clear results we will carry out MaAsLin3 analysis with genus counts.

## Aggregation

Use the function `microbiome::aggregate_taxa()` to produce a genus count based `phyloseq` object.

```{r, eval=FALSE}
#Aggregate ASV based phyloseq to genus phyloseq
genus_pseq <- microbiome::aggregate_taxa(pseq, "Genus", verbose=FALSE)
```

::: {.callout-note}
Biomarker detection/differential abundance allows you to determine if rare taxa are differentially abundant.
Due to this I would not reccomend aggregating rare taxa with `microbiome::aggregate_rare()` unless you have a good reason.
:::

Check the `phyloseq` object and number of genera.

```{r,eval=FALSE}
#Summarise
microbiome::summarize_phyloseq(genus_pseq)
#Number of genera
nrow(phyloseq::otu_table(genus_pseq))
```

## Clear up

We won't need the ASV phyloseq object anymore so you can remove it.

```{r, eval=FALSE}
#Remove pseq object
rm(pseq)
```