# Slope boxplots

On this page you'll:

- Create a boxplot visualisation of rarefaction slopes across the depth of samples.
- Create rarefaction slope boxplots side by side with depth boxplots
- Create a summary table showing the slopes and depths across the sample depths

## Slope box plot

Let's create the boxplots of rarefaction slopes across the sample depths.
The x-axis will be the slope values.
The y-axis will have the samples from lowest depth (bottom) to highest depth (top).

The purpose of this plot is to allow us to choose which sample we want to use as the minimum depth for rarefaction.
We can also use this for minimum depth sample removal.

```{r,eval=FALSE}
#Box plot of rare slopes
rareslope_boxplot <- rareslope_tbl |>
    ggplot2::ggplot(aes(y=RarefyDepthSample, x = RareSlope)) +
    ggplot2::geom_boxplot()
ggsave(filename = "rareslope_boxplot.png", plot = rareslope_boxplot, 
       device = "png", units="mm", width=200, height=500, dpi=300)
#Display plot
IRdisplay::display_png(file=
    "./rareslope_boxplot.png",
    width = 600)
```

In the plot you can see the range of rarefaction slopes starts high going up to a rarefaction slope of 0.15.
As we move up the plot the ranges reduce and the values decrease, a lower slope value the better.

It seems if we chose one of the rep4 samples as the rarefaction path we would still have quite high slopes.
Preferably we want the slopes to be lower than 0.01.

### Subset slope boxplot

Let's recreate the plot without the rep4 samples.
We'll definitely not want to use them and its hard to see the slope values on the x-axis of the other samples.

```{r, eval=FALSE}
#Box plot of rareslopes
rareslope_boxplot <- rareslope_tbl |>
    #Remove rows where "rep4" is in the RarefyDepthSample column
    dplyr::filter(!grepl("rep4",RarefyDepthSample)) |>
    #Plot
    ggplot2::ggplot(aes(y=RarefyDepthSample, x = RareSlope)) +
    ggplot2::geom_boxplot()
ggsave(filename = "rareslope_boxplot_no_rep4.png", plot = rareslope_boxplot, 
       device = "png", units="mm", width=200, height=500, dpi=300)
#Display plot
IRdisplay::display_png(file=
    "./rareslope_boxplot_no_rep4.png",
    width = 600)
```

In this case we can see that most of the rarefaction slopes are less than 0.01 if we used LD_ENV_rep5's depth as the rarefaction/minimum read depth size which is pretty good.

::: {.callout-tip title="Cookbook and reference links for Tidyverse function used above"}
- [`dplyr::filter()` with `grepl()`](https://neof-workshops.github.io/Tidyverse/dplyr/filter.html#regular-expression-matching-with-grepl)
:::

## Rarefaction slope and sample depth box plots

We can add the boxplot of sample of depths beside the slope box plots.
This is useful for determining a minimum read depth if you want to remove low depth samples.

::: {.callout-note}
This is not as useful for rarefaction size determination as all samples will have the same depth after rarefaction.
:::

To carry this out we need to "longify" the `SampleDepths` and `RareSlope` variables/columns so we can facet them.

```{R,eval=FALSE}
#Box plots of rareslopes and depth side by side
rareslope_boxplot <- rareslope_tbl |>
    #Longify Rareslope and Sampledepth columns
    tidyr::pivot_longer(
        c(SampleDepth,RareSlope), 
        names_to="metric",
        values_to="value") |>
    #Plot
    ggplot2::ggplot(aes(y=RarefyDepthSample, x = value)) +
    ggplot2::geom_boxplot() +
    #facet by metric column
    ggplot2::facet_grid(cols = dplyr::vars(metric), scales = "free_x")
ggsave(filename = "rareslope_and_depth_boxplot.png", plot = rareslope_boxplot, 
       device = "png", units="mm", width=200, height=500, dpi=300)
#Display plot
IRdisplay::display_png(file=
    "./rareslope_and_depth_boxplot.png",
    width = 600)
```

::: {.callout-tip title="Cookbook links for Tidyverse function used above"}
- [`tidyr::pivot_longer()`](https://neof-workshops.github.io/Tidyverse/tidyr/pivot_longer.html)
- ['ggplot2::facet_grid()'](https://neof-workshops.github.io/Tidyverse/ggplot2/facet.html#facet-grid)
:::

The right side box plots (SampleDepth) shows how the range of sample depths changes as samples are removed whn we move up the plot.
I.e. the second bottom boxplot (MD_KBC_rep4) shows the depth of all samples except for UD_KBC_rep4 (the bottom/lowest depth sample).

## Summary table

Finally we can create a summary table that will show the summary metrics of the box plots.
This will help us choose the rarefaction depth to choose based on the sample on the y-axis of our plots.

::: {.callout-note}
The summary plot will go from lowest depth (top) to highest depth (bottom).
:::

```{r,eval=FALSE}
#Summary output
summary_tbl <- rareslope_tbl |>
    dplyr::group_by(RarefyDepthSample) |>
    dplyr::summarise(
        rarefy_depth=mean(RarefyDepth),
        number_samples=n(),
        max_slope=max(RareSlope),
        mean_slope=mean(RareSlope),
        iqr_slope=IQR(RareSlope),
        min_slope=min(RareSlope),
        max_depth=max(SampleDepth),
        mean_depth=mean(SampleDepth),
        iqr_depth=IQR(SampleDepth),
        min_depth=min(SampleDepth)
    )
#View
summary_tbl
```

What if we used LD_ENV_rep5 as our cutoff sample for rarefaction/minimum sample depth filtering?

```{r,eval=FALSE}
#LD_ENV_rep5 info
summary_tbl |> dplyr::filter(RarefyDepthSample=="LD_ENV_rep5")
```

In this case if we we would have:

- A rarefaction size of 5,622
- 59 sample retained after rarefaction/minimum sample depth filtering
- Slope values ranging from 0.042 to 1.7e-05

This is fine for our data as we'll keep 5 replicates per site and media combination.
However, the rarefaction size is quite low and their are a lot of samples with depths >10,000.

What if we want to remove all the rep5 samples and use LD_CVP_rep1 (the lowest non rep5 or rep4 sample) as our sample threshold.

```{r,eval=FALSE}
#LD_CVP_rep1 info
summary_tbl |> dplyr::filter(RarefyDepthSample=="LD_CVP_rep1")
```

Now we have:

- Less samples (49), but we do have 4 replicates per combination which is still good
- A much higher rarefaction size of 10,433
- Better slope values ranging from 0.019 to 2.86e-12


::: {.callout-tip title="Cookbook links for Tidyverse function used above"}
- [`dplyr::group_by()`](https://neof-workshops.github.io/Tidyverse/dplyr/group_by.html)
- [`dplyr::summarise()`](https://neof-workshops.github.io/Tidyverse/dplyr/summarise.html)
:::